---
title: "foram_stat"
output: html_document
date: "2023-03-11"
---

Qs
0,1:First 10 Ma for average morphogroup richness and average MSI (linear logistic model and see if there is a difference with the first 5 ma as one category and the next 5 as another)
2: Separate out the high southern paleolatitudes, 50-65 degrees south (first 3 time bins, 10 cells) - south (blue) vs everything north (control for data numbers) - one tailed t-test
3: ESI and EPDI increase globally following the initiation of the Paleocene-Eocene Thermal Maximum (PETM, ~56 Ma), which continues until the termination of the Early Eocene Climatic Optimum (EECO, ~53.3-49.1 Ma) - t - test (three part - pre, during and post) - 62-56 Ma, 56-50 Ma, 50-44 Ma
4:Repeat the K-Pg southern hemisphere analysis for here (56-50 Ma) - 50-65 degrees south vs everything else (repeat for richness and ESI/EPDI)
5:Show a decrease in average morphogroup richness and average MSI across the Eocene-Oligocene boundary (linear model - logistic) and then test for how gradual the shift is (strength of logistic regression) (taken 34 Mya and plus minus 5 Mya so 39-29)
6:Repeat point 4 analysis for here (34-23 Ma) - 50-65 degrees south vs everything else (repeat for richness and MSI/MPDI)
7:Repeat 5 for oligocene-Miocene boundary (19 Ma as center) for MSI
QQ: Explore similarly for ~25Ma

```{r}
Q1=read.csv("data/Q1.csv")

y0=Q1$Mrich_1
y=scales::rescale(y0, to=c(0,1))
x=Q1$time
plot(y ~ x)
fit <- nls(y ~ SSlogis(x, Asym, xmid, scal), data = data.frame(x, y))

summary(fit)

sigmoid = function(params, x) {
  params[1] / (1 + exp(-params[3] * (x - params[2])))
}

params=coef(fit)
y2 <- sigmoid(params,x)
plot(y2,type="l", ylim=range(y))
points(y, pch=16)
abline(v=params[2], lty=2, col="red")
```
```{r}
y0=Q1$MSI_1
y=1-scales::rescale(y0, to=c(0,1))
x=Q1$time
plot(y ~ x)
fit <- nls(y ~ SSlogis(x, Asym, xmid, scal), data = data.frame(x, y))

summary(fit)

sigmoid = function(params, x) {
  params[1] / (1 + exp(-params[3] * (x - params[2])))
}

params=coef(fit)
    
y2 <- sigmoid(params,x)
plot(y2,type="l", ylim=rev(range(y)))
points(y, pch=16)
abline(v=params[2], lty=2, col="red")

#nagelkerke(model, null=null.model)
```



```{r}
Q5=read.csv("data/Q5.csv")

y0=Q5$Mrich_1
y=scales::rescale(y0, to=c(0,1))
x=Q5$time
plot(y ~ x)
fit <- nls(y ~ SSlogis(x, Asym, xmid, scal), data = data.frame(x, y))

summary(fit)

sigmoid = function(params, x) {
  params[1] / (1 + exp(-params[3] * (x - params[2])))
}

params=coef(fit)
    
y2 <- sigmoid(params,x)
plot(y2,type="l", ylim=range(y))
points(y, pch=16)
abline(v=params[2]+6, lty=2, col="red")
```
```{r}
Q5=read.csv("data/Q5.csv")

y0=Q5$MSI_1
y=scales::rescale(y0, to=c(0,1))
x=Q5$time
plot(y ~ x)
fit <- nls(y ~ SSlogis(x, Asym, xmid, scal), data = data.frame(x, y))

summary(fit)

sigmoid = function(params, x) {
  params[1] / (1 + exp(-params[3] * (x - params[2])))
}

params=coef(fit)
    
y2 <- sigmoid(params,x)
plot(y2,type="l", ylim=range(y))
points(y, pch=16)
abline(v=params[2]+6, lty=2, col="red")
```
```{r}
Q7=read.csv("data/66Ma/Qs/Q7.csv")

y0=Q7$Mrich_1
y=scales::rescale(y0, to=c(0,1))
x=Q7$time
plot(y ~ x)
fit <- nls(y ~ SSlogis(x, Asym, xmid, scal), data = data.frame(x, y))

summary(fit)

sigmoid = function(params, x) {
  params[1] / (1 + exp(-params[3] * (x - params[2])))
}

params=coef(fit)
    
y2 <- sigmoid(params,x)
plot(y2,type="l", ylim=range(y))
points(y, pch=16)
abline(v=params[2]+6, lty=2, col="red")
```

```{r}
Q7=read.csv("data/Q7.csv")

y0=1-Q7$MSI_1
y=scales::rescale(y0, to=c(0,1))
x=Q7$time
plot(y ~ x)
fit <- nls(y ~ SSlogis(x, Asym, xmid, scal), data = data.frame(x, y))

summary(fit)

sigmoid = function(params, x) {
  params[1] / (1 + exp(-params[3] * (x - params[2])))
}

params=coef(fit)
    
y2 <- sigmoid(params,x)
plot(y2,type="l", ylim = rev(range(y)))
points(y, pch=16)
abline(v=6+params[2], lty=2, col="red")
```

```{r}
Q0=read.csv("data/Q0.csv")
x=Q0$Time
y1=Q0$M_rich
y2=1-Q0$MSI
par(mar = c(5, 4, 4, 4) + 0.3)              # Additional space for second y-axis
plot(x, y1, pch = 16, col = 1, xlim = rev(range(x)), xlab="Time (Ma)", ylab="Avg. Morphotype richness")              # Create first plot
par(new = TRUE)                             # Add new plot
plot(x, y2, pch = 17,               # Create second plot without axes
     axes = FALSE, xlab = "", ylab = "", xlim = rev(range(x)), col="indianred")
axis(side = 4, at = pretty(range(y2)))      # Add second axis
mtext("1-Avg(MSI)", side = 4, line = 3,col="indianred")             # Add second axis label
```


```{r}
Q0=read.csv("data/Q0.csv")
x=Q0$Time
y1=Q0$E_rich
y2=1-Q0$ESI
par(mar = c(5, 4, 4, 4) + 0.3)              # Additional space for second y-axis
plot(x, y1, pch = 16, col = 1, xlim = rev(range(x)), xlab="Time (Ma)", ylab="Avg. Ecotype richness", col.lab = 1)              # Create first plot
par(new = TRUE)                             # Add new plot
plot(x, y2, pch = 17,               # Create second plot without axes
     axes = FALSE, xlab = "", ylab = "", xlim = rev(range(x)), col="goldenrod2")
axis(side = 4, at = pretty(range(y2)))      # Add second axis
mtext("1-Avg(ESI)", side = 4, line = 3, col="goldenrod2")             # Add second axis label
```


```{r}
#data_wide; Ecodiv
lm_mr_msi=NULL
for (i in 1:(length(roundage)-1)) {
  ZZ=c(summary(lm(MSI[,i]~Mrich[,i]))$coefficients[2,1],summary(lm(MSI[,i]~Mrich[,i]))$coefficients[2,4])
  lm_mr_msi=rbind(lm_mr_msi,ZZ)
}

#rownames(lm_mr_msi)= as.character(roundage)
colnames(lm_mr_msi)= c("coeff", "pval")

lm_mr_msi=as.data.frame(cbind(sort(roundage),lm_mr_msi))

plot(lm_mr_msi$coeff, pch=16, col="indianred", xlab="Time (Ma)", ylab="Linear model coefficient for latitudinal corr.", xlim = rev(range(x)))

```


```{r}
#data_wide; Ecodiv
lm_er_esi=NULL
for (i in 1:(length(roundage)-1)) {
  ZZ=c(summary(lm(ESI[,i]~Erich[,i]))$coefficients[2,1],summary(lm(ESI[,i]~Erich[,i]))$coefficients[2,4])
  lm_er_esi=rbind(lm_er_esi,ZZ)
}

#rownames(lm_mr_msi)= as.character(roundage)
colnames(lm_er_esi)= c("coeff", "pval")

lm_er_esi=as.data.frame(cbind(sort(roundage),lm_er_esi))

plot(lm_er_esi$coeff, pch=16, col="goldenrod2", xlab="Time (Ma)", ylab="Linear model coefficient for latitudinal corr.", xlim = rev(range(x)))

```

```{r}
QQ=read.csv("data/QQ.csv")

y0=QQ$MSI_1
y=scales::rescale(y0, to=c(0,1))
x=QQ$time
plot(y ~ x)
fit <- nls(y ~ SSlogis(x, Asym, xmid, scal), data = data.frame(x, y))

summary(fit)

sigmoid = function(params, x) {
  params[1] / (1 + exp(-params[3] * (x - params[2])))
}

params=coef(fit)
    
y2 <- sigmoid(params,x)
plot(y2,type="l", ylim=range(y))
points(y, pch=16)
abline(v=params[2], lty=2, col="red")
```


```{r}
QQ=read.csv("data/QQ.csv")

y0=QQ$Mrich_1
y=scales::rescale(y0, to=c(0,1))
x=QQ$time
plot(y ~ x)
fit <- nls(y ~ SSlogis(x, Asym, xmid, scal), data = data.frame(x, y))

summary(fit)

sigmoid = function(params, x) {
  params[1] / (1 + exp(-params[3] * (x - params[2])))
}

params=coef(fit)
    
y2 <- sigmoid(params,x)
plot(y2,type="l", ylim=range(y))
points(y, pch=16)
abline(v=params[2], lty=2, col="red")
```

